Spreading of a Facebook Application

### 2011-04-17 paper.tex
version for the course

\begin{abstract}
%% Text of abstract, 150 - 350 words
% The outcome was students' perceptions of preparedness.
Applications on Facebook are popular and attract many
people. Nevertheless, before installation people must acknowledge that an application
exists, and they must demand the application, too.
There may exist several reasons for people to install
applications, such as intensive usage within their social network or
personal interest in a special topic. Persuasion can
happen virally by Facebook's news stream, by marketing through
friends, or by directed search on 
Facebook. Differences in the spreading process are supposed to exist
for the different kinds of applications.

This paper investigates how a quiz, asking people five randomly
selected questions about dentistry health facts, spreads on
Facebook. The quiz is endowed with a social component: After
completion participants get more than the solution solely. Instead,
users can see images of friends who could answer the questions they got correctly
underneath the solution. In addition, they are told how good their
score was in comparison to their friends.

Analysis is supported by a tool maintaining an internal representation
of an extract of Facebook's social graph containing the participants
as nodes, their friendships as edges, and their basic information
retrieved from Facebook as well as quiz results as node
attributes. The tool allows for output of aspects of the internal graph
in DOT-language, and export of metrics to CSV, too.

Results show that participants who connect the network of existing
users to a new one often achieve a comparatively high score in the
quiz. In the most cases, new participants have strong relations to the network of people who
already participated. Indeed, the number of friends does not seem to
have any effect on whether a user gathers new participants for the quiz.
\end{abstract}

@@

\section{Introduction}
\label{sec:introduction}
Founded in 2004 Facebook\footnote{www.facebook.com} has
become a social network currently used by more than 600 million people
all over the world\cite{facebook500}. Since 2007, Facebook provides an API, permitting
third party developers to embed their own applications into
Facebook\footnote{http://developers.facebook.com/blog/archive}.
Subsequent to the F8 conference in April, 2010, Facebook simplifies
access to its social graph by allowing
external websites for integration in their social graph through Open
Graph Protocol.\cite{facebookStats}

Due to integration in Facebook's social graph, third party
applications not only spare user authentication procedures, but can also access
extracts of Facebook's social graph, too. These information detail a
user's friend list, basic information, and 
further data. Access to these data has to be granted by access tokens
\footnote{http://developers.facebook.com/docs/authentication/}. Thus,
applications let Facebook's social graph enlarge and make people
spend more time on Facebook. In addition, application developers
can spread their application virally in news
streams, and thereby they collect useful information about the social network of
their users.

@@

This paper investigates the way applications spread on
Facebook. For this purpose, a quiz was built asking users five randomly
selected questions about dentistry health facts. User data are
collected in a MySQL database. The Evaluation of the data is
accomplished through the output of a tool which maintains an internal
representation of the extract of Facebook's 
social graph the application is allowed to access.

@@

\section{Foundations}
\label{sec:background}
This section details concepts from Facebook, basics about the
\ac{dot} language, and finally, it defines the metrics the
data-evaluation tool (see section \ref{sec:methodology}) can put out. Facebook's
social graph concept is explained in 
section \ref{sec:social-graph}, section \ref{sec:integr-appl-with}
describes how applications can be integrated with Facebook, and
section \ref{sec:auth-auth} makes clear how applications can be
integrated with Facebook.

\subsection{Social Graph}
\label{sec:social-graph}
Facebook's \textit{social graph} can be regarded as its core. It
represents connections between people and things people care about 
\cite{facebookDev}. Since April, 2010, Facebook allows for access to its
social graph via the Open Graph Protocol\footnote{http://ogp.me/}, thus facilitating
external websites and third party applications to integrate in the social graph by becoming
additional nodes\cite{facebookDev2}. These additional nodes can for
instance represent entities such as movies, news articles,
applications, and much more. People on Facebook can add connections
from their user node to these additional nodes by clicking on plugins
provided by Facebook such as \textit{Like Buttons} or entering content in \textit{Comment
  Boxes}.  \cite{facebookDev2} As images of friends who liked an
entity, and comments of friends can be shown on external websites via
iframes, aspects from a user's social network are brought to these
websites, and for this reason, Facebook is said to make the Internet social.

In general, additional nodes representing external websites are integrated
unidirectional. Thus, Facebook retrieves the information in which
entities users were interested in, while the external websites do not
gain access to Facebook's social graph. Instead, the social
information is included through iframes.
In order to integrate nodes bi-directional, the user must permit the
access, and both, the user, and the entity, which wants to gain access
to the social graph, must be authenticated. This process is detailed in
section \ref{sec:auth-auth}. This happens in case of applications on
Facebook, thus they are included bi-directional and can grab
information from Facebook's social graph.

@@

In order to get access not only to public, but also to private
information, an access token has to be included in the Graph API
request. To get this token, the accessing entity must be
authorized and authenticated to do so. How this can be achieved is
explained in section \ref{sec:auth-auth}. When trying to access
non-public information without the needed access token, an error
is returned. Listing \ref{list:kbeckerfriends} gives an example for this.

@@

Besides the Graph API, Facebook's social graph can also be accessed
with the REST-API, which is in the processes of deprecating
\cite{facebookDev3}. Both, the Graph API and the REST-API support
O-Auth 2.0 for authentication and authorization. O-Auth 2.0 is
included in Facebook's \ac{sdks} (see section
\ref{sec:integr-appl-with} and \ref{sec:auth-auth}).

\subsection{Integration of Applications with Facebook}
\label{sec:integr-appl-with}
Applications on Facebook are hosted on external severs, and are loaded
into a \textit{Canvas Page} by using iFrames. After registration
(\url{http://www.facebook.com/developers/createapp.php}), applications
can be found under the URL \url{http://apps.facebook.com/your_app }.
As soon as the application exists for some month and has
some users registered, it can be added to Facebook's application
index.

Facebook provides several features to support the development
process. \ac{sdks} exist for popular programming languages and provide
sets of pre-defined functions for communication with Facebook. For
instance, they include the O-Auth 2.0 dialog which is described in
section \ref{sec:auth-auth}. In
addition, Facebook grants observation of its own development process by
making its raodmap public, and by showcasting its buglist. In addition, a
testserver exists to help developers to analyze if their application
is going to run with planned features from Facebook. 
Facebook provides an \textit{insights}-feature which makes
statistics about the applications available. \cite{oreilly}.

\subsection{Authentication and Authorization}
\label{sec:auth-auth}
In order to authorize and authenticate, Facebook recommends usage of
O-Auth 2.0 protocol by using the OAuth Dialog included in its
\ac{sdks}. As the application analyzed in this paper and described in section 
\ref{sec:method} implements the O-Auth Dialog from the PHP-\ac{sdk},
this variant will be described in detail. It shall just be mentioned,
that Facebook also allows for manual execution of the steps to receive
access tokens.\cite{facebookDevAuth}

The OAuth Dialog performs three steps, i.e. user authentication, 
and authorization as well as authentication of the
application. If these three steps are completed, the application is
issued an \textit{user access token}. If more than basic information
is needed from Facebook, scope parameters can be set.\cite{facebookDevAuth}

Figure \ref{fig:authorization} details how authorization and
authentication take place with OAuthDialog. The \textit{app server}
illustrates the server on which the app is actually
running. \textit{Facebook} represents all the Facebook servers, on
which perform authentication and authorization. The \textit{OAuth
  Dialog} runs on the browser of the client who is going to use the
application.

When clients access the canvas page of the application, they are
redirected to the OAuth Dialog (running on a Facebook server) and
provided with the application's ID. Then, user authentication and
app authorization is performed, by involving the client and Facebook
only. User authentication is done by logging in on Facebook (if necessary) and
authorization of the application is done by asking the user if he/she
wants to grant the application access to some data specified by the
application with the scope parameter. One should note, that no
login data are sent to the application, thus, this is how Facebook
establishes a trust relation. 
If the user authorizes the application, an \textit{authorization code} is returned and
the user is redirected to the application, to which the
\textit{authorization code} is forwarded. During registration of the
application, an \textit{application secret} code was
created. In order to authenticate the application, the
\textit{authorization code} and the \textit{application secret} have
to be sent to Facebook's \textit{Graph API Token Endpoint}. Facebook
then returns an \textit{access token}, which can be included in the
URL by which the Graph API can be accessed. However, Facebook's
\ac{sdks} also provide functions to perform calls to the Graph API,
and in this case, the access token is included automatically.

\subsection{Graphs with DOT language}
\label{sec:graph-repr-with}
The \ac{dot} language allows for plain text description of
graphs. There are few programs that can process \ac{dot} graphs, most
of them are part of the
\textit{Graphviz}\footnote{http://www.graphviz.org}-package. 
The language allows for directed and undirected graphs, attributes can
be set and colors as well as different shapes for nodes and edges can
be used. The basic syntax is very easy, the graph shown in figure
\ref{fig:dot} has the \ac{dot}-code given in listing \ref{list:dot}.

@@

Labels and colors can be added by defining the node in the head of the
graph description. An example is detailed in listing \ref{list:dot2}, and
the corresponding graph is given in figure \ref{fig:dot2}.

@@

\subsection{Software Metrics}
\label{sec:metrics}
Metrics define a mapping of a `particular measurable entity to a
numerical value' \cite{Lanz06a}. Thus, they allow for analysis and summarization
of properties which are oftentimes not measurable. In addition, they
facilitate detection of `outliers in large amounts of data'
\cite{Lanz06a}. Metrics can be plotted as diagrams or charts, or they
can be used for visualizations such as graphs connecting nodes
characterized by metrics. Since humans are trained in understanding signs and
pictures, visualizations supports understanding and identifying of
`hidden aspects'. \cite{Lanz06a} In this paper, they are
employed to characterize users of the quiz. Some metrics can be output
as \ac{pdf}, and others are used to create the graphs in \ac{dot} language.

The following metrics can be calculated:

\begin{itemize}
\item A user's number of friends.
\item A user's score.
\item The number of friends who also participated in the quiz.
\item The number of friends who used the application, and belong to a
new network, thus the only connector between the new users' nodes and
the other nodes of the graph is the node of the user analyzed.
\end{itemize}

\section{Methodology}
\label{sec:method}
This section describes the methodology. At first, the quiz which is
used to collect data from Facebook's social graph is described as well
as its implementation (section \ref{sec:quiz}).
Then, it is explained how data are collected for
further analysis and how they can be evaluated.

\subsection{The Analyzed Application}
\label{sec:quiz}
The application analyzed in this paper is a quiz. It is described in
\ref{sec:aboutQuiz}, and in \ref{sec:implmentation} its implementation
is detailed.

\subsubsection{About The Quiz}
\label{sec:aboutQuiz}
The quiz is about dental health fact and also challenging. It can be
found at the URL \url{http://apps.facebook.com/zahniquiz}.
Each participants is asked five questions randomly selected
from a set of twenty questions. A question has four answers, of
which either one, two, three or each answer can be correct, thus
the kind of answers refers to the term \textit{multiple selection}.
A question is scored to be answered correctly, if the correct subset
was selected by the user, thus each correct and no incorrect answer
were selected. Each question answered correctly scores one, thus the
highest score is five. Figure \ref{fig:quiz} gives a screenshot of the quiz.

The quiz is competitive. After the quiz is completed, users are told
their score, and in addition, they acknowledge how good they were in
comparison to other participants as well as to their 
friends. As shown in figure \ref{fig:solution}, for each question the
correct solution is given. A user can also invite friends to
participate. For this issue, profile pictures of a user's friends are
printed beyond the quiz. By clicking on such an image, an invitation
can be sent to the friend's Facebook profile.

\subsubsection{Implementation}
\label{sec:implementation}
The studied application makes use of 
Facebook's php-\ac{sdk} and JavaScript-\ac{sdk}\footnote{http://developers.facebook.com/docs/}. It requires
permissions (see \ref{sec:auth-auth}) for \textit{user$\_$birthday} and
\textit{publish$\_$stream}.
The questions and answers of the quiz are stored in a MySQL-database. PHP scripts
handle the creation of the quiz and
the Pear DB module \footnote{http://pear.php.net/} is used
for database queries.

The structure of the Facebook application is related to a
Model-view-controller architecture. \textit{Pear::DB} represents the
model, files stored in folder \textit{templates} act as view
(\textit{quizQuestion.php}, \textit{quiz.php},
\textit{quizResult.php}), and \textit{index.php} assumes the role of
the controller. \textit{library.php} contains a collection of
functions used by the various PHP scripts. \textit{facebook.php} and
scripts from \textit{facebook-sdk} folder are downloaded from
\url{https://github.com/facebook/php-sdk/} provide the functionality
of Facebook's PHP \ac{sdk}.

The view operates as follows: \textit{quiz.php} checks if the user
completed the quiz. If not, it prints the form, loads the styles and
includes \textit{quizQuestion.php} to print the questions. Otherwise, it
creates the page shown after the quiz is completed by printing a
statement commenting the user's score, and by providing users with 
the following options: Users can print their result into their news
stream and they can send invitations to their friends' Facebook
profiles. For both options, Facebook's JavaScript \ac{sdk} is
used. \textit{quizResult.php} contains the code to  
print the correct solution after the quiz is completed and is included
by \textit{quizQuestion.php}.

\paragraph{Database schema}
\label{sec:database-schema}
The database queries are performed with PHP
pear\footnote{http://pear.php.net/}-package. The following tables exist:
\begin{itemize}
\item \textbf{User} - entries are created for each user who participates as
well as for each of a user's friends. A flag details if user
participated or not.
\item \textbf{Friends} - creates entries for each friendship found in
users' friend lists.
\item \textbf{Questions} - contains the questions of the quiz.
\item \textbf{Answers} - contains the possible answers and a flag to mark them
as correct or false.
\item \textbf{GivenAnswers} - stores each answer given by a user.
\item \textbf{UserScore} - stores the score user had in the quiz.
\item \textbf{CorrectAnsweredQuestions} - stores which questions were correct
answered by which user, and which were not.
\end{itemize}

\paragraph{Application flow}
\label{sec:application-flow}
When a user accesses the canvas page of the quiz, 
the \textit{OAuth Dialog} is started automatically by the
\textit{controller}. If authorization and authentication are completed
successfully, the user's basic data as well as the birthday are stored together with
a timestamp in the \textit{User} table. Column \textit{has-used-app} is set to 1. In
addition, each user's friends list is also stored in \textup{User} and
\textit{Friends} table, and for these \textit{users}, columns
\textit{has-used-app} is set to 0.

% the quiz is created by randomly choosing
% five questions from the questions table and requiring the answers. To
% make the questions more attractive, the name of an image convenient to
% the questions topic is also stored in the questions table, and this
% image is placed on the left side of each question. Since multiple
% selection is possible, before the answers, which are also queried from
% the MySQL database, checkboxes are placed before each answer. 
Next, the quiz is created as mentioned in section
\ref{sec:implmenetation}.
After completion, its is evaluated the final score is computed. As
each answer ought to be stored and multiple selection of answers is
possible, user answers have to be stored in  a multidimensional.

The given answers are then stored in table \textit{GivenAnswers}. In
table (\textit{correctAnsweredQuestions)} it is stored which questions
were correct or incorrect answered. This happens to spare complicated
evaluations based on the \textit{GivenAnswers} table later on.
Finally, an average score is computed and stored in the user table.

@@

\subsection{Data evaluation}
\label{sec:data-evaluation}
For data evaluation, a tool coded in PHP exists. It allows for output
of metrics as \ac{csv} and graphs in \ac{dot}
language\footnote{http://www.graphviz.org/Documentation.php}.
The data evaluated by the tool are taken from
 database tables the quiz also makes use of.

As mentioned in previous sections, the tool maintains an extract of
Facebook's social graph internally. This graph contains users' basic
information, friendships, etc. The internal graph is
extended by information from the quiz, such as if the user has
installed the application, and which score a user had. The internal graph
consists of instances of class \textit{user}, which have references
among each other representing friendships. The whole set of users is
maintained by \textit{main.php}, which contains a list of all the user
instances.

Before the graph is created, user names can be made anonymous and the
results can be stored in the \textit{user} table to provide
reusability, too. User names are made anonymous by random selection of
a name from a list of popular baby names from 2009. The list of names was
taken from \url{http://www.ssa.gov/cgi-bin/popularnames.cgi}.   
For each gender, hundred names are stored in an array, and if the
gender is set, a name of same gender is chosen. Otherwise, the name is
taken randomly. The name of the person who spread the application
originally is always set to \textit{Alexandra}.

>>>>

\begin{abstract}
%% Text of abstract, 150 - 350 words
% The outcome was students' perceptions of preparedness.
Applications on Facebook are popular and attract many
people. Nevertheless, prior to installation people must acknowledge that an application
exists, and they must demand the application, too.
There may exist several reasons for people to install
applications, such as intensive usage within their social network or
personal interest in a special topic. Persuasion can
happen virally by Facebook's news stream or by marketing through
friends, but also through directed search on 
Facebook. Differences in the spreading process are supposed to exist
for the different kinds of applications.

This paper investigates how a quiz application spreads on
Facebook. A methodology is developed to conclude from a graph of
participants which people were very likely responsible for the application's spreading.

The quiz is about dentistry health facts and asks its participants five randomly
selected questions. In addition, it is endowed with a social component: After
completion of the quiz participants do not get the solution solely. Instead,
beyond the solution images of friends are displayed who knew the right answer of the
questions the user got. In addition, users are told how good their 
score was in comparison to their friends and to the whole set of users.

Analysis of the applications way of spreading is supported by a tool maintaining an internal representation
of an extract of Facebook's social graph. The tool's internal graph
represents the participants as nodes, their friendships as edges, and their basic information
retrieved from Facebook as well as quiz results as
attributes of the nodes. The methodology to identify people who
contributed to the spreading process was implemented, too. The tool
allows for output of aspects of the internal graph 
in DOT-language, and export of metrics to CSV, too.

Results show that participants who are supposed to contribute to the
application's spreading often achieve a comparatively high score in the
quiz. Particular user's posts in the news stream seem to influence the
applications growth, while other users post do not show this
effect. New participants often had strong relations to the network of people who
already participated. Indeed, the number of friends does not seem to
have any effect on whether a user gathers new participants for the quiz.
\end{abstract}

@@

\section{Introduction}
\label{sec:introduction}
Founded in 2004 Facebook\footnote{www.facebook.com on 14th of April 2011} has
become a social network currently used by more than 600 million people
all over the world\cite{facebook500}. Since 2007, Facebook provides an API, permitting
third party developers to embed their own applications into
Facebook.
Subsequent to the F8 conference in April, 2010, Facebook simplifies
access to its social graph by allowing
external websites for integration in their social graph through Open
Graph Protocol.\cite{facebookStats}

Due to integration in Facebook's social graph, third party
applications not only spare user from authentication procedures, but can also access
extracts of Facebook's social graph, too. This information details a
user's friend list, basic information, and 
further data. Access to this data has to be granted by access tokens. \cite{facebookDevAuth}
Thus, applications let Facebook's social graph enlarge and make people 
spend more time on Facebook. In addition, application developers
can spread their application virally in news
streams, and thereby they collect useful information about the social network of
their users.

@@

This paper investigates the way in which an application spreads on
Facebook. For this purpose, a quiz was built asking users five randomly
selected questions about dentistry health facts. User data are
collected in a MySQL database. The Evaluation of the data is
accomplished through the output of a tool which maintains an internal
representation of the extract of Facebook's 
social graph the application is allowed to access.

@@

\section{Foundations}
\label{sec:background}
This section details concepts from Facebook, basics about the
\ac{dot} language, and finally, it defines the metrics the
data-evaluation tool (see section \ref{sec:method}) can put out. Facebook's
social graph concept is explained in 
section \ref{sec:social-graph}, section \ref{sec:integr-appl-with}
describes how applications can be integrated with Facebook, and
section \ref{sec:auth-auth} illustrates the way how applications can be
integrated with Facebook.

\subsection{Facebook's Social Graph}
\label{sec:social-graph}
Facebook's \textit{social graph} can be regarded as its core. It
represents connections between people and things people care about 
\cite{facebookDev}. Since April 2010, Facebook allows access to its
social graph via the Open Graph Protocol\footnote{http://ogp.me/}, thus facilitating
external websites and third party applications to integrate in the social graph by becoming
additional nodes\cite{facebookDev2}. For instance, these additional
nodes can represent entities such as movies, news articles, 
applications, and many more. People on Facebook can add edges
between their user node to these additional nodes by clicking on plugins
provided by Facebook, such as \textit{Like Buttons} or entering content in \textit{Comment
  Boxes}.  \cite{facebookDev2} As images of friends liking an
entity, and comments of friends can be shown on external websites via
iframes, aspects from a user's social network are brought to these
websites. Therefore, Facebook is said to socialize the Internet.

In general, additional nodes representing external websites are integrated
unidirectionally. Thus, Facebook retrieves the information about the
entities users were interested in, whereas the external websites do not
gain access to Facebook's social graph. Instead, the social
information is included via iframes.
In order to integrate nodes bi-directionally, the user must permit 
access to its data, and both,  user and entity, wanting to gain access
to the social graph, must be authenticated. This process is detailed in
section \ref{sec:auth-auth} and matters in the case of applications on
Facebook. Thus, applications are included bi-directional and can grab
information from Facebook's social graph.

@@

In order to get access not only to public, but also to private
information, an access token has to be included in the Graph API
request. To get this token, the accessing entity has to be
authorized and authenticated to do so. A way to achieve this is
explained in section \ref{sec:auth-auth}. When trying to access
non-public information without the necessary access token, an error
is returned. Listing \ref{list:kbeckerfriends} gives an example.

@@

Besides the Graph API, Facebook's social graph can also be accessed
by using an older API that facebook refers to as its `REST-API'. This API is in the processes of deprecating
\cite{facebookDev3}. Both, the Graph API and the REST-API support
O-Auth 2.0 for authentication and authorization. O-Auth 2.0 is
included in Facebook's \ac{sdks} (see section
\ref{sec:integr-appl-with} and \ref{sec:auth-auth}).

\subsection{Integration of applications with Facebook}
\label{sec:integr-appl-with}
Applications on Facebook are hosted on external severs and are loaded
into a \textit{Canvas Page} by using iFrames. After registration
(\url{http://www.facebook.com/developers/createapp.php}), applications
can be found under the URL \url{http://apps.facebook.com/your_app }.
When the application has existed for some months and 
some users have registered themselves, it can be added to Facebook's application
index.

Facebook provides several features to support the development
process. \ac{sdks} exist for popular programming languages and provide
sets of pre-defined functions for the purpose of communication with Facebook. For
instance, they include the O-Auth 2.0 dialog which is described in
section \ref{sec:auth-auth}. In
addition, Facebook grants the observation of its own development process by
making its raodmap public and by publishing its buglist. Furthermore, a
testserver exists in order to help developers analyze whether their application
is going to run with planned features from Facebook. 
Facebook provides an \textit{Insights}-feature which makes
statistics about an application accessible. \cite{oreilly}.

\subsection{Authentication and authorization}
\label{sec:auth-auth}
In order to authorize and authenticate, Facebook recommends usage of
O-Auth 2.0 protocol by using the OAuth Dialog included in its
\ac{sdks}. As the application analyzed in this paper and described in section 
\ref{sec:method} implements the O-Auth Dialog from the PHP-\ac{sdk},
this variant will be described in detail. It ought be mentioned,
that Facebook also allows for manual execution of the steps to receive
access tokens.\cite{facebookDevAuth}  Indeed, this is
out of scope for the case of this paper.

The OAuth Dialog performs three steps, i.e. user authentication, 
and authorization as well as authentication of the
application. Whenever these three steps are completed, the application is
issued an \textit{user access token}. If more than basic information
is needed from Facebook, scope parameters can be set.\cite{facebookDevAuth}

Figure \ref{fig:authorization} details how authorization and
authentication take place with OAuthDialog. The \textit{app server}
depicts the server on which the application is actually
running. \textit{Facebook} represents all the Facebook servers, on
which authentication and authorization are performed. The \textit{OAuth
  Dialog} runs on the browser of the client who is going to use the
application.

When clients access the canvas page of the application, they are
redirected to the OAuth Dialog (running on a Facebook server) and
provided with the application's ID. Afterwards, user authentication and
authorization of the application are performed by involving the client and Facebook
only. User authentication is done by logging in on Facebook (if necessary) and
authorization of the application is done by asking the user whether he/she
wants to grant the application access to some data specified by the
application with the scope parameter. One should note, that no
login data are sent to the application. Thus, this is how Facebook
establishes a trust relation. 
If the user authorizes the application, an \textit{authorization code}
will be returned and
the user is redirected to the application to which the
\textit{authorization code} is forwarded. During the registration of the
application, an \textit{application secret} code is
created. In order to authenticate the application, the
\textit{authorization code} and the \textit{application secret} have
to be sent to Facebook's \textit{Graph API Token Endpoint}. Facebook
then returns an \textit{access token}, which can be included in the
URL by which the Graph API can be accessed. However, Facebook's
\ac{sdks} also provide functions to perform calls to the Graph API,
and in this case, the access token is included automatically.

\subsection{Graphs with DOT language}
\label{sec:graph-repr-with}
The \ac{dot} language allows for plain text description of
graphs. There are few programs that can process \ac{dot} graphs. Many
of them are part of the
\textit{Graphviz}\footnote{http://www.graphviz.org}-package. 
The language allows for creation of directed and undirected graphs. Attributes such
colors and shapes for nodes and edges can
be set. The graph shown in figure
\ref{fig:dot} has the \ac{dot}-code given in listing \ref{list:dot}.

@@

% Figure DOT 1 and 2
Labels and colors can be added by defining the node in the head of the
graph description. An example is detailed in listing \ref{list:dot2}, and
the corresponding graph is given in figure \ref{fig:dot2}.

@@

\subsection{Software Metrics}
\label{sec:metrics}
Metrics define a mapping from a `particular measurable entity to a
numerical value' \cite{Lanz06a}. Thus, they allow for analysis and summarization
of properties which oftentimes are not comparable without the metric. In addition, they
facilitate the detection of `outliers in large amounts of data'
\cite{Lanz06a}. Metrics can either be plotted as diagrams or charts, or they
can be used for visualizations. Such are graphs connecting nodes
characterized by metrics.

Since humans are trained in understanding signs and
pictures, visualizations support the understanding and identification of
`hidden aspects'.\cite{Lanz06a} In this paper, they are
employed to characterize users of the quiz. A tool evaluating the quiz
from the quiz exists, which can put out metrics
as \ac{csv}. In addition, it uses metrics are used to create the graphs in \ac{dot} language.

The following metrics can be printed to \ac{csv} files:

\begin{itemize}
\item A user's number of friends.
\item A user's score.
\item The number of friends who also participated in the quiz.
\item The number of friends who became participants due to
participation of the person analyzed. These persons are regarded as
persons who spread the application.
\end{itemize}

Further metrics are provided by Facebook Insights (see section
\ref{sec:integr-appl-with}). These include the number of Daily
App Installations, the Daily Application Usage, Monthly Active Users,
and the number of users who belong to a specific age group, or to a
specific gender.

\section{Methodology}
\label{sec:method}
In this section, it is detailed how the spreading of an application on
Facebook can be analyzed. At first, the quiz developed and employed for this issue is
explained. Afterwards, a method is described how the set of
participants can be transferred into a graph and how this
graph can be used to identify users who can have contributed to the applications
spreading. After this, the implementation of the quiz is detailed. In
the end of the chapter, it is described how the data were
evaluated. As a tool was developed for this purpose, a focus is put on
the way the tool  operates. The section finished with an explanation of
the way in which the spreading of the application was initialized.

\subsection{The application to be analyzed}
\label{sec:aboutQuiz}
The application to be analyzed is a quiz about dental health fact. It is to be
found at the URL \url{http://apps.facebook.com/zahniquiz}.
Each participant is asked five questions selected
from a set of twenty questions randomly. A question has four answers, of
which either one, two, three or each answer can be
correct. Consequently,
the kind of answers refers to the term \textit{multiple selection}.
A question is scored as a right answer, if the fitting subset was
selected by the user. 
Each question answered correctly scores one. Hence, the
highest score is five. Figure \ref{fig:quiz} gives a screenshot of the quiz.

The quiz is competitive. After the quiz is completed the score is
displayed, and in addition, they acknowledge how good they were in 
comparison to other participants as well as to their 
friends. As shown in figure \ref{fig:solution} for each question the
correct solution is shown.

To invite friends to the quiz, profile
pictures of a user's friends are 
printed beyond the quiz. By clicking on such an image,  requests to
participate in the applications 
can be sent to the appropriate friend.

\subsection{User-Graph and identification of people who can have contributed to the application's spreading}
\label{sec:user-graph-ident}
To identify characteristics of people who contributed to the spreading
of the application a graph can be created. In this section, this graph
will be termed User-Graph.

The nodes of the User-Graph represent either people who actually participated in the
quiz or friends of these people. Each node has some attributes: The
user's name, a randomly selected name to make the person anonymous,
and an id. If a person participated  
in the quiz, additional attributes are set. They refer to a
user's birthday, the gender, and information about the results the
user gained in the quiz. Edges are represented by friendships.

The graph can be used to identify people who can have contributed to
the application's spreading. A method to achieve this is shown in
figure \ref{fig:newNW}: User \textit{a} starts to use
the application and is friend of user \textit{b}, \textit{c}, and
\textit{d}. As user \textit{f}, and \textit{e} are only friends with
user \textit{b}, this user contributed to the application's spreading
very likely. It has to be mentioned that further ways exist to attract user \textit{e} or \textit{f}
to join the application. In this paper it is assumed that virality
is very important for the process of social transmission of an
application. Accepting this assumption, the existence of further ways
to attract people to participate becomes a weak argument.

Identification of people such as user \textit{b} can be achieved as
follows: At first,
the social network of the person who started to spread the application is
considered. For each participant, it is checked if participating
friends exist, who are not a friend of the person who started the
application's spreading. Those people are marked. Then, for each
marked person their friends are analyzed almost in the same way: For
each friend it is checked, if participating friends exist, not
included in the list considered list of friends and not included in
the list of friends of the person who started spreading the
applications. If people are found, they are marked, too. This
procedure is repeated as long as participating friends exist. In order
to avoid an infinite searching for people, for each marked person it
must be stored to which persons they probably spread the
application. If a person would be added twice, no further search is
needed within this part of the application.

It can be argued, that this method of identification does not consider
the shortest path on the graph. This is correct, but as it is not
known how the spreading took place it does not seem rational to
consider one path only. However, if more than one person is marked due to the
same participants, weights computed with respect to the distance
to the person who initialized the spreading could be used.

\subsection{Implementation}
\label{sec:implementation}
The implementation of the quiz was done using PHP, Java Script, and MySQL.
Facebook's PHP-\ac{sdk} and
JavaScript-\ac{sdk}\footnote{http://developers.facebook.com/docs/}
were employed as well. The application requires
permissions for \textit{user$\_$birthday} and
\textit{publish$\_$stream} (see \ref{sec:auth-auth}). These
permissions permit the application to retrieve a user's birthday and to
send posts to a user's news stream.

The questions and answers of the quiz are stored in a MySQL-database. PHP scripts
handle the creation of the quiz and
the Pear DB module \footnote{http://pear.php.net/} is used
for database queries.

The structure of the Facebook application is related to a
Model-view-controller architecture. \textit{Pear::DB} represents the
model, files stored in folder \textit{templates} act as view
(\textit{quizQuestion.php}, \textit{quiz.php},
\textit{quizResult.php}), and \textit{index.php} assumes the role of
the controller. \textit{library.php} contains a collection of
functions used by the various PHP scripts. \textit{facebook.php} and
scripts from the \textit{facebook-sdk} folder are downloaded from
\url{https://github.com/facebook/php-sdk/} provide the functionality
of Facebook's PHP \ac{sdk}.

The view operates as follows: \textit{quiz.php} checks if the user
completed the quiz, and subject to this, either the quiz is shown or
the pages containing the solution are displayed and the user is
provided with
the following options:  The result can be printed into the news
stream and invitations can be send to friends' Facebook
profiles. For both options Facebook's JavaScript \ac{sdk} is
used. \textit{quizResult.php} contains the code to  
print the correct solution after the quiz is completed and is included
by \textit{quizQuestion.php}.

\subsubsection{Database schema}
\label{sec:database-schema}
The database queries are performed with the 
Pear\footnote{http://pear.php.net/} library. The following tables exist:
\begin{itemize}
\item \textbf{User} - entries are created for each participating user as
well as for each of a user's friends. A flag details if a user
participated or not.
\item \textbf{Friends} - creates entries for each friendship found in
users' list of friends.
\item \textbf{Questions} - contains the questions of the quiz.
\item \textbf{Answers} - contains the possible answers and a flag to mark them
either right or wrong.
\item \textbf{GivenAnswers} - stores each answer given by a user.
\item \textbf{UserScore} - stores the score a user had in the quiz.
\item \textbf{CorrectAnsweredQuestions} - stores which questions were
answered right by which user, and which were not.
\end{itemize}

\subsubsection{Application flow}
\label{sec:application-flow}
When a user accesses the canvas page of the quiz, first
the \textit{OAuth Dialog} is executed. 
If authorization and authentication are completed
successfully, the user's basic data and the birthday are stored together with
a timestamp in the \textit{User} table. Column \textit{has-used-app}
is set to 1 if a user participated. In
addition each user's list of table is also stored in \textup{User} and
\textit{Friends} table and for these users column
\textit{has-used-app} is set to 0.

% the quiz is created by randomly choosing
% five questions from the questions table and requiring the answers. To
% make the questions more attractive, the name of an image convenient to
% the questions topic is also stored in the questions table, and this
% image is placed on the left side of each question. Since multiple
% selection is possible, before the answers, which are also queried from
% the MySQL database, checkboxes are placed before each answer. 
Afterwards the quiz is created as mentioned in section
\ref{sec:implementation}.
After completion, the quiz is evaluated and the final score is computed. Because
each answer ought to be stored and multiple answers are
possible, answers of users are stored in  a multidimensional array.

The given answers are stored in table \textit{GivenAnswers}. In
table \textit{correctAnsweredQuestions} it is stored which questions
were answered right or wrong. This happens to spare complicated
evaluations based on the \textit{GivenAnswers} table later on.
Finally, an average score is computed and stored in the user table.

@@

\subsection{Data evaluation}
\label{sec:data-evaluation}
To evaluate the data retrieved from Facebook and collected when users
participated in the quiz  a tool was developed. It is coded in PHP and
employs MySQL as well as the Pear library to access the
data which were gathered when people participated in the quiz. The
tool allows for output of metrics as \ac{csv} and graphs in \ac{dot}
language\footnote{http://www.graphviz.org/Documentation.php}. These
metrics are used to identify characteristics of people stored in
the \textit{User} table. The graphs are used to visualize the whole
process of the application's spreading. 
Besides usage of the tool, data can also be retrieved using database
queries. Such queries were performed to get the number of people in the user table and
the number of people who actually participated.

\subsubsection{Implementation of the tool for evaluation}
\label{sec:impl-tool-eval}
As mentioned in previous sections, the tool maintains an extract of
Facebook's social graph internally. This graph contains users' basic
information, friendships, and so forth. The internal graph is
extended by data gained through quiz evaluations. This data includes
information about application usage and the score the users
received. The nodes and edges of the internal graph 
are implemented as instances of class \textit{user}. Edges on the
graph represent friendships. These are represented by lists of
references to the appropriate user objects and stored as a user object
attribute.
The whole set of user objects is
maintained by \textit{main.php}, where they are stored as a list.

Before the graph is created, user names can be made anonymous and the
results can be stored in the \textit{user} table to provide
reusability. User names are made anonymous by random selection of
a name from a list of popular baby names from 2009. The list of names was
taken from \url{http://www.ssa.gov/cgi-bin/popularnames.cgi}.   
For each gender hundred names are stored in an array and selected
randomly depending on the user's gender. 
For users whose gender is not known, name and sex are set randomly. 
The name of the person who started spreading the application
originally is set to \textit{Alexandra} by default.

The tool identifies people who can have contributed to the
application's spreading as follows: People who are supposed to have
spread the quiz are found by recursively checking if an user of 
the application has participating friends whose friends also participated
in the quiz, but are neither contained in the analyzed user's list of
friends nor in the list of friends of the person
who spread the application in the beginning. Then, references to the
people to which the user is supposed to have spread the application
are stored in a list. If a reference is added twice, the recursion is stopped.
